Spatial and temporal decay of SARS-CoV-2 environmental contamination of the healthcare environment

### 2021-12-08 manuscript.Rmd
Mark edits.

### Acknowledgements:  
This research, BJK, MJZ, and EL are supported in part by a CDC Cooperative Agreement FOA#CK16-004 "Epicenters for the Prevention of Healthcare Associated Infections”. BJK is supported by the National Institute for Allergy and Infectious Diseases (K23 AI121485 and L30 AI120149). MJZ is supported by the National Institute for Allergy and Infectious Diseases (K23 AI143925).

@@

### Word Count:    
?????

@@

* __Results__: The probability of detecting SARS-CoV-2 RNA in a patient room did not vary with distance. However, we found that surface type predicted probability of detection, with floors and high-touch surfaces having the highest probability of detection (floors odds ratio (OR) `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_site_categoryFloor") %>% pull(Estimate_exp) %>% signif(.,3)` (95% CrI `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_site_categoryFloor") %>% pull(Q2.5_exp) %>% signif(.,3)` to `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_site_categoryFloor") %>% pull(Q97.5_exp) %>% signif(.,3)`); high-touch elevated OR `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_high_touchTRUE") %>% pull(Estimate_exp) %>% signif(.,3)` (95% CrI `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_high_touchTRUE") %>% pull(Q2.5_exp) %>% signif(.,3)` to `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_high_touchTRUE") %>% pull(Q97.5_exp) %>% signif(.,3)`)). Patients requiring high-flow oxygen, positive airway pressure, or mechanical ventilation were associated with increased surface contamination (OR `r m_combined_summary %>% filter(model_name == "COVID-19 Severity") %>% filter(grepl("SignificantOxygen",param)) %>% pull(Estimate_exp) %>% signif(.,3)` (95% CrI `r m_combined_summary %>% filter(model_name == "COVID-19 Severity") %>% filter(grepl("SignificantOxygen",param)) %>% pull(Q2.5_exp) %>% signif(.,3)` to `r m_combined_summary %>% filter(model_name == "COVID-19 Severity") %>% filter(grepl("SignificantOxygen",param)) %>% pull(Q97.5_exp) %>% signif(.,3)`)). The probability of surface elevated contamination decayed with prolonged hospitalization, but the probability of floor detection increased with duration of the local pandemic wave.

@@

Despite infection control measures, transmission of SARS-CoV-2 infection has been observed in the hospital and other healthcare settings. In a recent study of SARS-CoV-2-infected and -uninfected patients inadvertently roomed together, 39% of exposed, previously uninfected patients developed COVID-19. Risk of SARS-CoV-2 transmission was associated with source patients who had low cycle threshold (Ct), i.e., high viral load, in their diagnostic testing [@Karan2021-cm].

>>>>

### Acknowledgements:  
This research, BJK, MJZ, and EL are supported in part by a CDC Cooperative Agreement FOA#CK16-004 "Epicenters for the Prevention of Healthcare Associated Infections”. BJK is supported by the National Institute for Allergy and Infectious Diseases (K23 AI121485 and L30 AI120149). MJZ is supported by the National Institute for Allergy and Infectious Diseases (K23 AI143925). We thank the Bushman Laboratory at the University of Pennsylvania for quantitative PCR methods support.

@@

### Word Count:    
- Abstract: 199  
- Body: 2227

@@

* __Results__: The probability of detecting SARS-CoV-2 RNA in a patient room did not vary with distance. However, we found that surface type predicted probability of detection, with floors and high-touch surfaces having the highest probability of detection (floors odds ratio (OR) `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_site_categoryFloor") %>% pull(Estimate_exp) %>% signif(.,3)` (95% CrI `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_site_categoryFloor") %>% pull(Q2.5_exp) %>% signif(.,3)` to `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_site_categoryFloor") %>% pull(Q97.5_exp) %>% signif(.,3)`); high-touch elevated OR `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_high_touchTRUE") %>% pull(Estimate_exp) %>% signif(.,3)` (95% CrI `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_high_touchTRUE") %>% pull(Q2.5_exp) %>% signif(.,3)` to `r m_combined_summary %>% filter(model_name == "Distance Fixed" & param == "b_high_touchTRUE") %>% pull(Q97.5_exp) %>% signif(.,3)`)). <mark>Increased surface contamination was observed in room where patients required high-flow oxygen, positive airway pressure, or mechanical ventilation</mark> (OR `r m_combined_summary %>% filter(model_name == "COVID-19 Severity") %>% filter(grepl("SignificantOxygen",param)) %>% pull(Estimate_exp) %>% signif(.,3)` (95% CrI `r m_combined_summary %>% filter(model_name == "COVID-19 Severity") %>% filter(grepl("SignificantOxygen",param)) %>% pull(Q2.5_exp) %>% signif(.,3)` to `r m_combined_summary %>% filter(model_name == "COVID-19 Severity") %>% filter(grepl("SignificantOxygen",param)) %>% pull(Q97.5_exp) %>% signif(.,3)`)). The probability of elevated surface contamination decayed with prolonged hospitalization, but the probability of floor detection increased with duration of the local pandemic wave.

@@

Despite infection control measures, transmission of SARS-CoV-2 infection has been observed in the hospital and other healthcare settings. In a recent study of SARS-CoV-2-infected and -uninfected patients inadvertently roomed together, 39% of exposed, previously uninfected patients developed COVID-19. Risk of SARS-CoV-2 transmission was associated with source patients who had low cycle threshold (Ct), i.e., high viral load, in their diagnostic testing, <mark>with 91.7% of those exposed to such source patients developing infection</mark>[@Karan2021-cm].
